Key Derivation From Noisy Sources With More Errors Than Entropy

### 2014-06-10 trunk/constantRate/ConstCompFuzzExt.tex
Adding info-theoretic construction and reusability

@@

\section{Reusable Fuzzy Extractors}
\label{sec:reusable}

Fuzzy extractors should authenticate users to multiple different services/devices.  That is, instead of creating a single pair $(r, p)$ from an initial reading $w$, we read the noisy sources multiple times~($w_1,..., w_q$) and produce $(r_1, p_1),..., (r_q, p_q)$.  This problem is known as a reusable fuzzy extractor~\cite{Boyen2004}.  There are two basic versions of the problem, first where the adversary sees $p_1,..., p_q$~(outsider security), and second where the adversary also sees some subset of the derived keys as well~(insider security).  Ideally, we want $w_1,..., w_q$ and the unseen keys to be ``as good'' as in the single use case.  Boyen constructs a reusable cryptographic fuzzy extractor based on the code-offset or the syndrome sketches from~\cite{JW99, DBLP:journals/siamcomp/DodisORS08} instantiated with linear codes.  For simplicity here, we define the insider security and note any scheme that achieves insider security also achieves outsider security.

There are two primary questions about the reusability of a fuzzy extractor.  The first is how often it can be reused, the second is what correlation is allowed between $w_1,..., w_q$.  Boyen~\cite{Boyen2004} considers fuzzy extractors that are arbitrarily reusable.  Their construction is secure when the correlations between $w_1,.., w_q$ come from a subgroup of permutations used in the construction of the fuzzy extractor.  They show in the setting of arbitrary reuse such a strong restriction on correlation is necessary~\cite[Theorem 11]{Boyen2004}.  In particular, they show how any information-theoretic fuzzy extractor secure with arbitrary reuse can be transformed into their construction.  However, there is no reason to expect that noise between repeated readings comes from a well-behaved subgroup of permutations.

However, their negative result is inherently information-theoretic, the adversary must exhaust all randomness of the fuzzy extractor and computes statistical distance over exponential size strings.  There are two natural restrictions to the definition of reusability: 1) limit the number of reuses of the fuzzy extractor 2) provide security against computational adversaries.  The second restriction implies the adversary will only observe a polynomial number of invocations of the fuzzy extractor.fuzzy extractor will only be used a polynomial number of times.

Instead of restricting the type of correlations, we limit the adversary to be computationally bounded~(and thus sees only a limited number of reuses).
We present a simplified non-adaptive definition and note that our results apply equally well to the adaptive definition.

\begin{definition}[Insider Reusable Fuzzy Extractors]
\label{def:reusable fuzz}
Let $\mathcal{W}$ be a family of distributions over $\mathcal{M}$.  Let $(\gen, \rep)$ be a $(\mathcal{M}, \mathcal{W}, \kappa, t)$-computational fuzzy extractor that is $(\epsilon, s)$-hard with error $\delta$.  Let $W_1,..., W_q$ be a joint distribution such that $\forall i, W_i\in\mathcal{W}$.  We say that $(\gen, \rep)$ is $q$-reusable if for all $i$
\begin{align*}
\delta^{\mathcal{D}_{s_{sec}}}(&(R_1, P_1), ..., (R_i, P_i),..., (R_q, P_q), \\
&(R_1, P_1),..., (U_\kappa, P_i),..., (R_q, P_q)) < \epsilon.
\end{align*}
\end{definition}

\consref{cons:sampling} is the fuzzy extractor that allows for arbitrary correlation between $w_1,..., w_q$ as long as each $w_i$ comes from an admissible distribution.  %Our construction can be made secure information-theoretically for fixed values of $q$ or computationally for an arbitrary value $q = \poly(n)$.  Our construction is based on a modification of the computational fuzzy extractor of Canetti et al.~\cite[Construction 5.2]{canetti2014key}.  Their construction is secure for a restricted class of distributions.\footnote{This is necessary as they secure distributions with more errors than entropy.  In order to provide any meaningful security guarantee, the source distribution must be restricted.  See the introduction of~\cite{canetti2014key} for discussion.}  In addition to showing the reusability of their construction, we provide two significant improvements:

\begin{theorem}

\end{theorem}

@@

\section{Proof of \lemref{lem:cond and cext}}
\label{sec:cond and cext}
\lnote{I need to check this}
\begin{proof}
It suffices to show if there is some distinguisher $D'$ of size $s'$ where
\[\delta^{D'}((\cext(X; U_d), U_d, P'), (U_\kappa, U_d, P'))>\epsilon_{cond}+ \epsilon_{ext}\]
 then there is an distinguisher $D$ of size $s_{cond}$ such that for all $Y$ with $\Hav(Y|P') \geq \tilde{m}$,
 \[
 \delta^{D}((X,P'), (Y, P'))\geq \epsilon_{cond}.
 \]
Let $D'$ be such a distinguisher.  That is,
\[
\delta^{D'}(\cext(X, U_d)\times U_d \times P', U_\kappa\times U_d\times P')> \epsilon_{ext}+\epsilon_{cond}.
\]
Then define $D$ as follows.  On input $(y, p')$ sample $seed\leftarrow U_d$, compute $r\leftarrow \cext(y; seed)$ and output $D(r, seed, p')$.  Note that $|D| \approx s' + |\cext| +d= s_{cond}$.  Then we have the following:
\begin{align*}
\delta^{D}((X, P'), (Y, P'))&= \delta^{D'}((\cext(X, U_d), U_d, P'), \cext(Y, U_d), U_d, P')\\
&\geq \delta^{D'}((\cext(X, U_d), U_d, P'), (U_\kappa\times U_d \times P')) \\
&- \delta^{D'}((U_\kappa\times U_d \times P'), (\cext(Y, U_d), U_d, P'))\\
&>\epsilon_{cond}+\epsilon_{ext}- \epsilon_{ext} = \epsilon_{cond}.
\end{align*}
Where the last line follows by noting that $D'$ is of size at most $s_{ext}$.  Thus $D$ distinguishes $X$ from all $Y$ with sufficient conditional min-entropy.  This is a contradiction.
\end{proof}

>>>>

\subsection{Reusable Fuzzy Extractors}
\label{sec:reusable}

Ideally, a fuzzy extractor allows a user to authenticate to multiple different services/devices.  That is, instead of creating a single pair $(r, p)$ from an initial reading $w$, we read the noisy source multiple times~($w_1,..., w_q$) and produce $(r_1, p_1),..., (r_q, p_q)$.  This problem is known as a reusable fuzzy extractor~\cite{Boyen2004}.  There are two basic versions of the problem, first where the adversary sees $p_1,..., p_q$~(outsider security), and second where the adversary also sees some subset of the derived keys as well~(insider security).  Ideally, we want $w_1,..., w_q$ and the unseen keys to be ``as good'' as in the single use case.  Boyen constructs a reusable cryptographic fuzzy extractor based on the code-offset or the syndrome sketches from~\cite{JW99, DBLP:journals/siamcomp/DodisORS08} instantiated with linear codes.  For simplicity here, we define the insider security and note any scheme that achieves insider security also achieves outsider security.

There are two primary questions about the reusability of a fuzzy extractor.  The first is how often it can be reused, the second is what correlation is allowed between $w_1,..., w_q$.  Boyen~\cite{Boyen2004} considers fuzzy extractors that are reusable an unbounded number of times.  Their construction is secure when the correlations between $w_1,.., w_q$ come from a subgroup of permutations used in the construction of the fuzzy extractor.  They show in the setting of unbounded reuse such a strong restriction on correlation is necessary~\cite[Theorem 11]{Boyen2004}.  In particular, they show how any information-theoretic fuzzy extractor secure with unbounded reuse can be transformed into their construction.  However, there is no reason to expect that noise between repeated readings comes from a well-behaved subgroup of permutations.

However, their negative result is inherently information-theoretic, the adversary must exhaust all randomness of the fuzzy extractor and computes statistical distance over exponential size strings.  There are two natural restrictions to the definition of reusability: 1) limit the number of reuses of the fuzzy extractor 2) provide security against computational adversaries.  The second restriction implies the adversary will only observe a polynomial number of invocations of the fuzzy extractor.

We limit the adversary to be computationally bounded~(and thus sees only a limited number of reuses).
We present a non-adaptive definition and note that our results extend to the adaptive definition presented by Boyen~\cite[Definition 7]{Boyen2004}.\bnote{We need to come back to justifying this or provide a proof that our scheme works in the adaptive setting.}

\begin{definition}[Insider Reusable Fuzzy Extractors]
\label{def:reusable fuzz}
Let $\mathcal{W}$ be a family of distributions over $\mathcal{M}$.  Let $(\gen, \rep)$ be a $(\mathcal{M}, \mathcal{W}, \kappa, t)$-computational fuzzy extractor that is $(\epsilon, s)$-hard with error $\delta$.  Let $W_1,..., W_q$ be a joint distribution such that $\forall i, W_i\in\mathcal{W}$.  Let $(R_i, P_i)$ be the distribution formed by running $\gen$ on the distribution $W_i$.  We say that $(\gen, \rep)$ is $q$-reusable if for all $i$
\begin{align*}
\delta^{\mathcal{D}_{s}}(&(R_1, P_1), ..., (R_i, P_i),..., (R_q, P_q), \\
&(R_1, P_1),..., (U_\kappa, P_i),..., (R_q, P_q)) < \epsilon.
\end{align*}
\end{definition}

%\consref{cons:sampling} is the fuzzy extractor that allows for arbitrary correlation between $w_1,..., w_q$ as long as each $w_i$ comes from an admissible distribution.  %Our construction can be made secure information-theoretically for fixed values of $q$ or computationally for an arbitrary value $q = \poly(n)$.  Our construction is based on a modification of the computational fuzzy extractor of Canetti et al.~\cite[Construction 5.2]{canetti2014key}.  Their construction is secure for a restricted class of distributions.\footnote{This is necessary as they secure distributions with more errors than entropy.  In order to provide any meaningful security guarantee, the source distribution must be restricted.  See the introduction of~\cite{canetti2014key} for discussion.}  In addition to showing the reusability of their construction, we provide two significant improvements:

@@

@@

\section{An information-theoretic construction with $\Huse\le 0$}
In this section we show an information-theoretic analogue to \consref{cons:first construction}.  %This construction can essentially be seen as replacing the computational point obfuscation with information-theoretic randomness extractors.  
Using this weaker object requires us to make a stronger assumption on our starting distribution.  Furthermore, the important features of \consref{cons:sampling}, $\Huse \le 0$ for a small alphabet and reusability, crucially rely on computational assumptions.  Achieving these benefits information-theoretically is an open question.  We will construct an information-theoretic fuzzy conductor which can be converted to a fuzzy extractor using a randomness extractor~(using the information-theoretic analogue of \lemref{lem:cond and cext}).

The main idea of this construction is to first condense entropy from each block of the source and then apply a standard secure sketch to the extracted information.  Intuitively, we are then able from a large alphabet while paying entropy loss on the smaller alphabet.  %Alternatively, this construction can simply be seen as replacing the computational point function obfuscation with an information-theoretic point-function obfuscation~(this is only possible for high enough entropy levels).
We first review some information-theoretic definitions.

\begin{definition}
\label{def:secure sketch}
An $(\mathcal{M},m, \tilde{m}, t)$-\emph{secure sketch} with error $\delta$ is a pair of randomized procedures, ``sketch'' $(\sketch)$ and ``recover'' $(\rec)$, with the following properties:
\begin{enumerate}
\item The sketching procedure \sketch on input $w\in\mathcal{M}$ returns a bit string $s\in\{0,1\}^*$.
\item The recovery procedure \rec takes an element $w'\in\mathcal{M}$ and a bit string $s\in\{0,1\}^*$.  The \emph{correctness} property of secure sketches guarantees that if $\dis(w,w')\leq t$, then $\Pr[\rec(w',\sketch(w))=w]\geq 1-\delta$ where the probability is taken over the coins of $\sketch$ and $\rec$.
\item The \emph{security} property guarantees that for any distribution $W$ over $\mathcal{M}$ with min-entropy $m$, $\Hav(W|\sketch(W))\geq \tilde{m}$.
\end{enumerate}
A secure sketch is \emph{efficient} if \sketch and \rec run in expected polynomial time. 
\end{definition}

\begin{definition}
\label{def:conductor}
A function $\cond : \mathcal{Z}\times \zo^d\rightarrow \mathcal{Y}$ is a $(m, \tilde{m}, \epsilon)$-average case randomness condenser if whenever $\Hoo(W)\ge m$, then there exists a distribution $Y$ with $\Hoo(Y)\ge \tilde{m}$ and $(\cond(W, seed), seed) \approx_\epsilon (Y, seed)$.
\end{definition}

Note that randomness condensers can be built significantly more efficiently than randomness extractors~(see for example~\cite{dodis2013key}).

\begin{definition}
\label{def:fuzzy conductor}
A pair of randomized procedures ``generate'' ($\gen$) and ``reproduce'' ($\rep$) is an $(\mathcal{M}, \mathcal{W}, \tilde{m}, t$)-fuzzy conductor that is $\epsilon$-hard with error $\delta$ if $\gen$ and $\rep$ satisfy \defref{def:comp fuzzy extractor}, except the last condition is replaced with the following condition:
\begin{itemize}
\item For any distribution $W\in \mathcal{W}$, the string $r$ has is close to a distribution with average min-entropy conditioned on $P$.  That is there exists a distribution $Y$ with $\Hav(Y | P)\ge \tilde{m}$ such that $(R ,P)\approx_\epsilon (Y, P)$.
\end{itemize}
\end{definition}

\begin{lemma}
Let $\mathcal{W}$ be the family of $(\alpha, \beta)$-partial block sources over $\mathcal{Z}^\ell$ and let $\cond: \mathcal{Z} \times \zo^d\rightarrow \mathcal{Y}$ be a $(\alpha, \tilde{\alpha}, \epsilon)$-average case randomness conductor.  Let $(\sketch, \rec)$ be $(\mathcal{Y}^\ell, \tilde{\alpha}(\ell - \beta), \tilde{m}, t)$-secure sketch with error $\delta$.  Then the following is a $(\mathcal{Z}^\ell, \mathcal{W}, \tilde{m}, t)$-fuzzy conductor that is $\epsilon \ell$ hard with error $\delta$:
\begin{center}
\begin{tabular}{c|c}
\begin{minipage}{3in}
\textbf{\gen}
\begin{enumerate}
\item \underline{Input}: $w = w_1,..., w_\ell$
\item For $j=1,..., \ell$:
\begin{enumerate}[(i)]
\item Sample $seed_i\leftarrow \zo^d$.
\item Set $v_i = \cond(w_i, seed_i)$.
\end{enumerate}
\item Set $ss \leftarrow \sketch(v_1,..., v_\ell)$.
\item Set $p = (ss, seed_1,..., seed_\ell)$.
\item Output $(v, p)$.
\end{enumerate}
 \end{minipage} &
\begin{minipage}{3in}
\textbf{\rep}
\begin{enumerate}
\item \underline{Input}: $(w', p = (ss, seed_1,..., seed_\ell))$
\item For $j=1,..., \ell$:
\begin{enumerate}[(i)]
\item Set $v_i = \cond(w_i, seed_i)$.
\end{enumerate}
\item Output $v = \rec(v', ss)$.
\end{enumerate}
\vspace{0.7in}
\end{minipage}
\end{tabular}
\end{center}
\end{lemma}
\begin{proof}
Let $W\in \mathcal{W}$.
It suffices to argue correctness and security.  We first argue correctness.  
When $w_i = w_i'$, then $\cond(w_i , seed_i) = \cond(w_i', seed_i)$ and thus $v_i = v_i'$.  Thus, for all $w, w'$ where $\dis(w, w')\le t$, then $\dis (v, v')\le t$.  Then by correctness of $(\sketch, \rep)$, $\Pr[\rec(v', \sketch(v)) = v]\ge 1-\delta$.

We now argue security.  Denote by $seed$ the random variable consisting of all $\ell$ seeds and $V$ the entire string of generated $V_1,..., V_\ell$.
Our goal is to show that $\Hav(V | p)$ is close to a distribution with average min-entropy $\tilde{m}$.  We first argue that there exists a distribution $Y$ where $\Hav(Y | seed)\ge \tilde{\alpha}\beta$ and $(V, seed_1,..., seed_\ell)\approx (Y, seed_1,.., seed_\ell)$.  First note since $W$ is $(\alpha, \beta)$-partial block distribution that
there exists a set of indices $J$ where $|J| \geq \ell - \beta$ such that the following holds:
\[
\forall j\in J, \forall w_1,..., w_\gamma \in W_1,..., W_\gamma, \Hoo(W_j | W_1 = w_1,..., W_{j-1}=w_{j-1}) \geq \alpha.
\]
Then consider the first element of $j_1\in J$, $\forall w_1,..., w_{j_1-1}\in W_1,..., W_{j_1-1}$, 
\[\Hoo(W_{j_1} | W_1 = w_1,..., W_{j_1-1} = w_{j_1-1})\ge \alpha.\]

%\[(\ext (W_1, seed_1), seed_1) \approx_{\epsilon} (U_{\mathcal{Y}}, seed_1)\]
%Then since $W$ is a block-source, 
%\[ \forall w_1\in W_1, \Hoo(W_2 | w_1)\ge k.\]
\noindent 
Thus, there exists a distribution $Y_{j_1}$ with $\Hav(Y_{j_1} | seed_{j_1}) \ge \tilde{\alpha}$ such that 
\[(\cond (W_{j_1}, seed_{j_1}), seed_{j_1}, W_1,..., W_{j_1-}) \approx_{\epsilon} (Y_{j_1}, seed_{j_1}, W_1,..., W_{j_1})\]
and since $(seed_1,..., seed_{j_1})$ are independent of these values
\[(\cond (W_{j_1},seed_{j_1}), W_{j_1-1},..., W_1, seed_{j_1}, ..., seed_{1}) \approx_{\epsilon} (Y_{j_1}, W_{j_1-1},..., W_1, seed_{j_1}, , ...,  seed_{1})\]
consider the random variable $Z_{j_1} =( Y_{j_1}, \cond(W_{j_1-1},seed_{j_1-1}),..., \cond(W_{1}, seed_{1}))$ and note that \[\Hav(Z_{j_1} | seed_1,...,seed_{j_1})\ge \alpha'.\]
Applying a deterministic function does not increase statistical distance and thus,
\begin{align*}
(\cond (W_{j_1}, seed_{j_1}), \cond(W_{j_1-1}, seed_{j_1-1}),..., \cond(W_1, seed_1), seed_{j_1},..., seed_{1}) \\\approx_{\epsilon} (Z_{j_1}, seed_{j_1},..., seed_1)
\end{align*}

\noindent
By a hybrid argument there exists a distribution $Z$ with $\Hav(Z | seed) \ge \alpha(\ell -\beta)$ where 
\[
(\cond(W_\ell, seed_\ell), ..., \cond(W_1, seed_1), seed_\ell,..., seed_1) \approx_{\epsilon\ell} (Z, seed_\ell,...,  seed_1).\]
By the security of $(\sketch, \rec)$ we know that $\Hav(Z | seed, ss) \ge \tilde{m}$.  %Note that $U_{\mathcal{Y}}^\ell$ is independent of $seed_1,..., seed_\ell$ and so $\Hav(U_{\mathcal{Y}}^\ell, seed_1,..., seed_\ell, p)\ge \tilde{m}$. 
That is, 
\[
(V, seed_1,..., seed_\ell, p) \approx_{\epsilon\ell} (U_{\mathcal{Y}}^\ell, seed_1,..., seed_\ell, p).\]  This completes the proof.
\end{proof}

\subsection{Discussion of Information-Theoretic Construction}
\bnote{write this}
\section{Proof of \lemref{lem:cond and cext}}
\label{sec:cond and cext}
\lnote{I need to check this}
\begin{proof}
It suffices to show if there is some distinguisher $D'$ of size $s'$ where
\[\delta^{D'}((\cext(X; U_d), U_d, P'), (U_\kappa, U_d, P'))>\epsilon_{cond}+ \epsilon_{ext}\]
 then there is an distinguisher $D$ of size $s_{cond}$ such that for all $Y$ with $\Hav(Y|P') \geq \tilde{m}$,
 \[
 \delta^{D}((X,P'), (Y, P'))\geq \epsilon_{cond}.
 \]
Let $D'$ be such a distinguisher.  That is,
\[
\delta^{D'}(\cext(X, U_d)\times U_d \times P', U_\kappa\times U_d\times P')> \epsilon_{ext}+\epsilon_{cond}.
\]
Then define $D$ as follows.  On input $(y, p')$ sample $seed\leftarrow U_d$, compute $r\leftarrow \cext(y; seed)$ and output $D(r, seed, p')$.  Note that $|D| \approx s' + |\cext| +d= s_{cond}$.  Then we have the following:
\begin{align*}
\delta^{D}((X, P'), (Y, P'))&= \delta^{D'}((\cext(X, U_d), U_d, P'), \cext(Y, U_d), U_d, P')\\
&\geq \delta^{D'}((\cext(X, U_d), U_d, P'), (U_\kappa\times U_d \times P')) \\
&- \delta^{D'}((U_\kappa\times U_d \times P'), (\cext(Y, U_d), U_d, P'))\\
&>\epsilon_{cond}+\epsilon_{ext}- \epsilon_{ext} = \epsilon_{cond}.
\end{align*}
Where the last line follows by noting that $D'$ is of size at most $s_{ext}$.  Thus $D$ distinguishes $X$ from all $Y$ with sufficient conditional min-entropy.  This is a contradiction.
\end{proof}
